humanCdGF_predict_cfu

### 2016-10-14 code/HumanCdGF_predict_cfu.Rmd
dual y axis for partial dependence plots

heat.plot <- function(input_maxtrix, high_color = '#660000', low_color = '#FFFFFF', legend_title = NA) {
  long.df <- data.frame(input_maxtrix)
  colnames(long.df) <- c(1:10)
  long.df$community_day <- as.character(c(0:10))
  long.df %>% gather(cfu_day, variable, -community_day) %>% 
    ggplot(aes(cfu_day, community_day)) + 
    geom_tile(aes(fill = variable), color = 'white') +
    scale_fill_gradient(low = low_color, high = high_color, name = legend_title, limits = c(0,1)) +
    labs(x = 'Day (CFU)', y = 'Day (Community)') + 
    theme(axis.line.x = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank()) + 
    scale_x_discrete(expand = c(0, 0), limits = c('1','2','3','4','5','6','7','8','9','10')) + 
    scale_y_discrete(expand = c(0, 0), limits = rev(c('0','1','2','3','4','5','6','7','8','9','10'))) + 
    theme(legend.position = 'top', legend.key.width = unit(2, 'cm'))
}
```

```{r functions}
rsq <- matrix(NA, nrow = 11, ncol = 10)
mse <- matrix(NA, nrow = 11, ncol = 10)
otu_votes <- rep(0, ncol(rel_abund_shared_01))
names(otu_votes) <- colnames(rel_abund_shared_01)
importance.df <- c()
for(m in 0:10){
  sample_list <- shared_list_days %>% 
    filter(day == m) %>% 
    select(-day)
  community.df <- rel_abund_shared_01 %>% 
    right_join(sample_list) %>% 
    select(-sample_id)
    for(n in 1:10){
      cfu_shared.df <- meta_file %>% 
        filter(day == n) %>% 
        select(mouse_id, log_cfu) %>% 
        right_join(community.df) %>% 
        select(-mouse_id) %>% 
        filter(!is.na(log_cfu))

@@

```{r top otus, fig.width=6, fig.height=6}
heat.plot(rsq_n, '#003300', legend_title = expression(paste('R'^'2')))
#plot(otu_votes_n[otu_votes_n >= 1], xlab = 'OTU', ylab = 'Freq')
```

@@

```{r fig.height=5, fig.width=8}
meta_data %>% 
  filter(!is.na(day), !cage_id == 'NP1', 
         Early_Euth == FALSE, cdiff_strain == 431) %>% 
  ggplot(aes(x = factor(day), y = log_cfu, color = cage_id)) + 
    geom_line(aes(group = mouse_id), size=.25, alpha=0.4) + geom_point() +
    stat_summary(fun.y=mean, geom="line", aes(group = cage_id)) + 
    theme_bw() + labs(x = 'Day', y = 'C Diff CFU (log10)') +
    theme(legend.justification=c(1,0), legend.position=c(1,0), 
          legend.title=element_blank())
```

@@

```{r identify reason for lowered R2, fig.height=5, fig.width=8}
log_cutoff <- 2.5

@@

```{r fig.width=6,fig.height=6}
heat.plot(rsq_r, '#003300', legend_title = expression(paste('R'^'2')))
```

@@

```{r plot features,fig.width=8, fig.height=20}
meta_data %>% 
  filter(!sample_id == 'OUT2-2533-D8', log_cfu >= log_cutoff) %>% 
  filter(!cage_id == 'NP1', Early_Euth == FALSE, cdiff_strain == 431) %>% 
  full_join(rel_abund_shared_01) %>% 
  select(day, sample_id, mouse_id, cage_id, 
         one_of(names(otu_votes_r[otu_votes_r >= 1]))) %>% 
  gather(otu, rel_abund, Otu00002:Otu00250) %>% 
  filter(!is.na(day)) %>% 
  ggplot(aes(x=factor(day),y=rel_abund, color = cage_id)) +
    geom_line(aes(group = mouse_id), size=.25, alpha=0.4) + geom_point() +
    stat_summary(fun.y=mean, geom="line", aes(group = cage_id)) + 
    facet_grid(otu ~., scales='free_y') +
    theme_bw() + labs(x = 'Day', y = 'Relative Abundance (%)')
```

@@

```{r predict median CFU heatmap, fig.width=6,fig.height=6}

@@

cage_color <- rainbow(length(unique(rf_prediction$cage_id)))
for (i in seq_along(importance_final)) {
  partial_otu <- 
    partialPlot(rf_n_features, top_colonized_df, names(importance_final)[i],
                xlab='', main=paste(tax_label[i]), 
                ylim = c(min(top_colonized_df$log_cfu_med),max(top_colonized_df$log_cfu_med)))
  for(m in seq_along(unique(rf_prediction$cage_id))){
    cage <- unique(rf_prediction$cage_id)[m]
    points(y = top_colonized_df[rf_prediction$cage_id %in% cage, 'log_cfu_med'],
         x = top_colonized_df[rf_prediction$cage_id %in% cage , names(importance_final)[i]], 
         pch = 21, bg = cage_color[m])
  }
}

@@

cage_color <- rainbow(length(unique(rf_cr_prediction$cage_id)))
for (i in seq_along(importance_final)) {
  partial_otu <- 
    partialPlot(rf_col_rate_features, top_col_rate_df, names(importance_final)[i],
                xlab='', main=paste(tax_label[i]), 
                ylim = c(min(top_col_rate_df$ratio),max(top_col_rate_df$ratio)))
  for(m in seq_along(unique(rf_cr_prediction$cage_id))){
    cage <- unique(rf_cr_prediction$cage_id)[m]
    points(y = top_col_rate_df[rf_cr_prediction$cage_id %in% cage, 'ratio'],
         x = top_col_rate_df[rf_cr_prediction$cage_id %in% cage , names(importance_final)[i]], 
         pch = 21, bg = cage_color[m])
  }
}

>>>>

```{r functions, eval = F}

heat.plot <- function(input_maxtrix, high_color = '#660000', low_color = '#FFFFFF', legend_title = NA) {
  long.df <- data.frame(input_maxtrix)
  colnames(long.df) <- c(1:10)
  long.df$community_day <- as.character(c(0:10))
  long.df %>% gather(cfu_day, variable, -community_day) %>% 
    ggplot(aes(cfu_day, community_day)) + 
    geom_tile(aes(fill = variable), color = 'white') +
    scale_fill_gradient(low = low_color, high = high_color, name = legend_title, limits = c(0,1)) +
    labs(x = 'Day (CFU)', y = 'Day (Community)') + 
    theme(axis.line.x = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank()) + 
    scale_x_discrete(expand = c(0, 0), limits = c('1','2','3','4','5','6','7','8','9','10')) + 
    scale_y_discrete(expand = c(0, 0), limits = rev(c('0','1','2','3','4','5','6','7','8','9','10'))) + 
    theme(legend.position = 'top', legend.key.width = unit(2, 'cm'))
}

rsq <- matrix(NA, nrow = 11, ncol = 10)
mse <- matrix(NA, nrow = 11, ncol = 10)
otu_votes <- rep(0, ncol(rel_abund_shared_01))
names(otu_votes) <- colnames(rel_abund_shared_01)
importance.df <- c()
for(m in 0:10){
  sample_list <- shared_list_days %>% 
    filter(day == m) %>% 
    select(-day)
  community.df <- rel_abund_shared_01 %>% 
    right_join(sample_list) %>% 
    select(-sample_id)
    for(n in 1:10){
      cfu_shared.df <- meta_file %>% 
        filter(day == n) %>% 
        select(mouse_id, log_cfu) %>% 
        right_join(community.df) %>% 
        select(-mouse_id) %>% 
        filter(!is.na(log_cfu))

@@

```{r top otus, fig.width=6, fig.height=6, eval = F}
heat.plot(rsq_n, '#003300', legend_title = expression(paste('R'^'2')))
#plot(otu_votes_n[otu_votes_n >= 1], xlab = 'OTU', ylab = 'Freq')
```

@@

```{r fig.height=5, fig.width=8, eval = F}
meta_data %>% 
  filter(!is.na(day), !cage_id == 'NP1', 
         Early_Euth == FALSE, cdiff_strain == 431) %>% 
  ggplot(aes(x = factor(day), y = log_cfu, color = cage_id)) + 
    geom_line(aes(group = mouse_id), size=.25, alpha=0.4) + geom_point() +
    stat_summary(fun.y=mean, geom="line", aes(group = cage_id)) + 
    theme_bw() + labs(x = 'Day', y = 'C Diff CFU (log10)') +
    theme(legend.justification=c(1,0), legend.position=c(1,0), 
          legend.title=element_blank())
```

@@

```{r identify reason for lowered R2, fig.height=5, fig.width=8, eval = F}
log_cutoff <- 2.5

@@

```{r fig.width=6,fig.height=6, eval = F}
heat.plot(rsq_r, '#003300', legend_title = expression(paste('R'^'2')))
```

@@

```{r plot features,fig.width=8, fig.height=20, eval = F}
meta_data %>% 
  filter(!sample_id == 'OUT2-2533-D8', log_cfu >= log_cutoff) %>% 
  filter(!cage_id == 'NP1', Early_Euth == FALSE, cdiff_strain == 431) %>% 
  full_join(rel_abund_shared_01) %>% 
  select(day, sample_id, mouse_id, cage_id, 
         one_of(names(otu_votes_r[otu_votes_r >= 1]))) %>% 
  gather(otu, rel_abund, Otu00002:Otu00250) %>% 
  filter(!is.na(day)) %>% 
  ggplot(aes(x=factor(day),y=rel_abund, color = cage_id)) +
    geom_line(aes(group = mouse_id), size=.25, alpha=0.4) + geom_point() +
    stat_summary(fun.y=mean, geom="line", aes(group = cage_id)) + 
    facet_grid(otu ~., scales='free_y') +
    theme_bw() + labs(x = 'Day', y = 'Relative Abundance (%)')
```

@@

```{r predict median CFU heatmap, fig.width=6,fig.height=6, eval = F}

@@

cage_color <- rainbow(length(unique(rf_prediction$cage_id)))
for (i in seq_along(importance_final)) {
    par(mar = c(5,5,2,5))
    partialPlot(rf_n_features, top_colonized_df, names(importance_final)[i],
                xlab='', main=paste(tax_label[i]))
    par(new = T)
    plot(y = top_colonized_df[ , 'log_cfu_med'],
         x = top_colonized_df[ , names(importance_final)[i]], 
         pch = 21, bg = 'gray', xlab = NA, ylab = NA, axes = F)
    axis(side = 4)
    mtext(side = 4, line = 3, 'Log10 CFU')
# To color points by cage - Change pch to 21 and un-comment following lines
#  for(m in seq_along(unique(rf_prediction$cage_id))){
#    cage <- unique(rf_prediction$cage_id)[m]
#    points(y = top_colonized_df[rf_prediction$cage_id %in% cage, 'log_cfu_med'],
#         x = top_colonized_df[rf_prediction$cage_id %in% cage , names(importance_final)[i]], 
#         pch = 21, bg = cage_color[m], xlab = NA, ylab = NA, axes = F)
#  }
}

@@

cage_color <- rainbow(length(unique(rf_cr_prediction$cage_id)))
for (i in seq_along(importance_final)) {
    partialPlot(rf_col_rate_features, top_col_rate_df, names(importance_final)[i],
                xlab='', main=paste(tax_label[i]))
    par(new = T)
    plot(y = top_col_rate_df[ , 'ratio'],
         x = top_col_rate_df[ , names(importance_final)[i]], 
         pch = 21, bg = 'gray', xlab = NA, ylab = NA, axes = F)
    axis(side = 4)
    mtext(side = 4, line = 3, 'Relative Colonization')
# To color points by cage - Change pch to 21 and un-comment following lines
#  for(m in seq_along(unique(rf_cr_prediction$cage_id))){
#    cage <- unique(rf_cr_prediction$cage_id)[m]
#    points(y = top_col_rate_df[rf_cr_prediction$cage_id %in% cage, 'ratio'],
#         x = top_col_rate_df[rf_cr_prediction$cage_id %in% cage , names(importance_final)[i]], 
#         pch = 21, bg = cage_color[m])
#  }
}
```
